---
title: "Rmd with googlesheets4"
subtitle: "Deploying to RStudio Connect"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```


This repository contains an example of an R Markdown document that reads from googlesheets and it deployed to RStudio Connect. Deploying content that interacts with a Google API can be a rather cumbersome process. Google authorization is usually done interactively via the [gargle](https://gargle.r-lib.org/) package. However, when you need to deploy to a headless machine this isn't feasible. 

This example illustrates using an deploying an R Markdown document which uses data from googlesheets. The data in the Rmd is accessed from a private googlesheets. Then it is deployed and scheduled on RStudio Connect. This process is outline in great detail [here](https://gargle.r-lib.org/articles/non-interactive-auth.html#sidebar-1-deployment).

## Authenticating 

Load the [`googlesheets4`](https://googlesheets4.tidyverse.org/) package. 

```{r}
library(googlesheets4)
```

By default gargle obfuscates the storage of the authentication token. Here we sepcify a project level directory `.secrets` which will contains our Google token. We will set the `garge_oath_cache` option to refer to this `.secrets` directory. We can check where the token will be cached with `gargle::gargle_oath_cache()`.


```{r}
# designate project-specific cache
options(gargle_oauth_cache = ".secrets")

# check the value of the option, if you like
gargle::gargle_oauth_cache()
```

Next we will have to preform the interactive authentication just once. Doing this will generate the token and store it for us. You will be required to select an email account to authenticate with. 

```{r}
# trigger auth on purpose --> store a token in the specified cache
# a broswer will be opened
googlesheets4::sheets_auth()
```

Now that you have completed the authentication and returned to R, we can double check that the token was cached in `.secrets`.

```{r}
# see your token file in the cache, if you like
list.files(".secrets/")
```

Voila! Let's deauthorize in our session so we can try authenticating once more, but this time without interactivity. 

```{r}
# deauth
sheets_deauth()
```

In `sheets_auth()` we can specify where the token is cached and which email we used to authenticate.

```{r}
# sheets reauth with specified token and email address
sheets_auth(
  cache = ".secrets",
  email = "josiah@email.com"
  )
```


Alternatively, we can specify these in the `options()` and run the authentication without an arguments supplied. Let's first deauth in our session to try authenticating again.

```{r}
# deauth again
sheets_deauth()

# set values in options
options(
  gargle_oauth_cache = ".secrets",
  gargle_oauth_email = "josiah@email.com"
)

# run sheets auth
sheets_auth()
```

Now that we are sure that authorization works without an interactive browser session, we should migrate the options into an `.Rprofile` file. This way, when an R session is spun up the options will be set from session start. Meaning, if you use `sheets_auth()` within your R Markdown document it will knit without having to open the browser.

## Deploying to Connect

In order for the deployment to RStudio Connect to work, the `.secrets` directory and `.Rprofile` files need to be in the bundle. Be sure to do this from the `Add Files` button. If you cannot see the files because they are hidden from Finder you cran press `cmnd + shift + .`. Then publish!


